
---
title: "Decoding Intracellular Pathogen of H3N2 at the Single-Cell level using SCKIT"
author: "Wei Zhang"
date: "2021/09/10"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: no
    theme: cerulean
    highlight: tango
    number_sections: yes
    df_print: tibble
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[UTF-8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, eval = TRUE,
                      warnings = FALSE, message = FALSE,
                      fig.width = 6, fig.height = 5)
```

Taking the in-vitro experiment of H3N2 infection data (SRA Accession number: [SRP239555](https://www.ncbi.nlm.nih.gov/sra/?term=SRP239555)) as an example, we used [PathogenTrack](https://github.com/rstatistics/PathogenTrack) to identify H3N2 infected cells at the single-cell level and used [SCKIT](https://github.com/rstatistics/SCKIT) to analyze and explore the biological functions that may be related to H3N2 infection.

# Install SCKIT from GitHub
```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!requireNamespace("devtools", quietly = TRUE))
    BiocManager::install("devtools")
if (!requireNamespace("Yeskit", quietly = TRUE))
    devtools::install_github("ncrna/Yeskit")
```

# Import SCKIT
```{r, warning=FALSE}
library(Yeskit)
library(topGO)
```

# Importation
```{r fig.width=4, fig.height=4}
Bystander <- scRead(sample_name = "Bystander", data_dir = system.file("extdata/H3N2_10X_matrix/Bystander/", package="Yeskit"), gene_column = 2, project_name = "H3N2", group_name = "Bystander", meta_file = system.file("extdata/H3N2_10X_matrix/Bystander/microbes.tsv", package="Yeskit"))
Infected <- scRead(sample_name = "Infected", data_dir = system.file("extdata/H3N2_10X_matrix/Infected/", package="Yeskit"), gene_column = 2, project_name = "H3N2", group_name = "Infected", meta_file = system.file("extdata/H3N2_10X_matrix/Infected/microbes.tsv", package="Yeskit"))
```

# Integration
```{r}
H3N2_integrated <- scIntegrate(object.list=list(Bystander, Infected), object.names = c("Bystander", "Infected"), batch.rm = "harmony", res = 0.7)
```

# FindMarkers
```{r}
slot(H3N2_integrated, "misc")$Markers <- Seurat::FindAllMarkers(object = H3N2_integrated, assay = 'RNA', test.use = 'MAST')
```

# Differential analysis
## Differential analysis between Infected and Bystander
```{r}
slot(H3N2_integrated, "misc")$Infected_vs_Bystander <- scDGE(object = H3N2_integrated, comparison = c("Infected", "Bystander"), group.by = "group", min.cells = 10, logFC = 0.25, clusters = NULL)
```

## Differential analysis between H3N2_positive and H3N2_negative 
```{r}
slot(H3N2_integrated, "misc")$H3N2 <- scPathogenDGE(object = H3N2_integrated, species.by = "H3N2", min.cells = 5)
```

# GO annotation
## GO annotation of Markers
```{r}
slot(H3N2_integrated, "misc")$Markers.GO <- scGO(object = H3N2_integrated, key = "Markers", logFC = 0.25, only.pos = TRUE, reference = "human")
```

## GO annotation of DGEs between Infected and Bystander
```{r}
slot(H3N2_integrated, "misc")$Infected_vs_Bystander.GO <- scGO(object = H3N2_integrated, key = "Infected_vs_Bystander", logFC = 0.25, only.pos = FALSE, reference = "human")
```

## GO annotation of DGEs between H3N2_positive and H3N2_negative 
```{r}
slot(H3N2_integrated, "misc")$H3N2.GO <- scPathogenGO(object = H3N2_integrated, key = "H3N2", clusters = NULL, species = "H3N2", logFC = 0.25)
```

# MSigDB scoring
```{r}
H3N2_integrated <- scMsigdbScoring(object = H3N2_integrated, category = "H", geneSets = NULL)
```

# Visualization
## Visualization of cell clusters by scDimPlot
```{r, fig.height=4, fig.width=8, fig.align="center"}
scDimPlot(object = H3N2_integrated, reduction = "umap", cols = NULL, split.by = "sample", ncol = 2, pt.size = 2)
```

## Visualization of cell densities by scDensityPlot
```{r, fig.height=4, fig.width=8, fig.align="center"}
scDensityPlot(object = H3N2_integrated, reduction = "umap", split.by = "sample", ncol = 2)
```

## Visualization of cell population fractions by scPopulationPlot, the x axis stands for clusters
```{r, fig.height=4, fig.width=5, fig.align="center"}
scPopulationPlot(object = H3N2_integrated, by = "cluster", cols = "sc", order = c("Bystander", "Infected"))
```

## Visualization of cell population fractions by scPopulationPlot, the x axis stands for samples
```{r, fig.height=4, fig.width=4, fig.align="center"}
scPopulationPlot(object = H3N2_integrated, by = "sample", order = c("Bystander", "Infected"))
```

## Visualization of meta data by scVizMeta
```{r, fig.height=4, fig.width=4, fig.align="center"}
scVizMeta(object = H3N2_integrated, reduction = "umap", signature="H3N2", title = "H3N2", raster = TRUE, split.by = "sample", pt.size = 2, interval = c(Abundant = 1000, Large = 500, Medium = 100, Small = 10, Single = 1, None = 0))
```

## Visualization of H3N2-infected cell fractions by scPathogenRatioPlot
```{r, fig.height=4, fig.width=4, fig.align="center"}
scPathogenRatioPlot(object = H3N2_integrated, species = "H3N2", split.by = "sample", ncol = 2)
```

## Visualization of DGEs by scVolcanoPlot
```{r, fig.height=4, fig.width=4, fig.align="center"}
scVolcanoPlot(H3N2_integrated, key = "Infected_vs_Bystander", cluster = "0", top_n = 10)
```

## Visualization of enriched GO terms for up-regulated genes by scGOBarPlot
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGOBarPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", top_n = 10, direction = "up", cluster = "0")
```

## Visualization of enriched GO terms for down-regulated genes by scGOBarPlot
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGOBarPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", top_n = 10, direction = "down", cluster = "0")
```

## Visualization of enriched GO terms for up-regulated genes by scGODotPlot
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGODotPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", direction = "up", top_n = 10, font.size = 10)
```

## Visualization of enriched GO terms for down-regulated genes by scGODotPlot
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGODotPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", direction = "down", top_n = 10, font.size = 10)
```

## Visualization of up-regulated GO terms in cluster-0
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGODotPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", direction = "up", top_n = 6, font.size = 10, order.genes = FALSE, clusters = "0")
```

## Visualization of down-regulated GO terms in cluster-0
```{r, fig.height=2, fig.width=3, fig.align="center"}
scGODotPlot(object = H3N2_integrated, key = "Infected_vs_Bystander.GO", ont = "BP", direction = "down", top_n = 10, font.size = 10, order.genes = FALSE, clusters = "0")
```

## Visualization of HALLMARK_INFLAMMATORY_RESPONSE pathway
```{r, fig.height=2, fig.width=5, fig.align="center"}
scScoreDimPlot(H3N2_integrated, signature = "HALLMARK_INFLAMMATORY_RESPONSE", split.by="sample", ncol = 2, pt.size = 2)
```

## Visualization of HALLMARK_TNFA_SIGNALING_VIA_NFKB pathway
```{r, fig.height=2, fig.width=5, fig.align="center"}
scScoreDimPlot(H3N2_integrated, signature = "HALLMARK_TNFA_SIGNALING_VIA_NFKB", split.by="sample", ncol = 2, pt.size = 2)
```

## Visualization of HALLMARK_APOPTOSIS pathway
```{r, fig.height=2, fig.width=5, fig.align="center"}
scScoreDimPlot(H3N2_integrated, signature = "HALLMARK_APOPTOSIS", split.by="sample", ncol = 2, pt.size = 2)
```

# Session Information
```{r, echo=FALSE}
sessionInfo()
```

